---
title: "Create a frequency plot of interpolated trajectory points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
This vignette demonstrates how to use the get_traj() function to create a frequency plot of trajectory points. 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(testtraj)
library(terra)
```
Before you start, make sure that the package folder contains the hyts_std executable file. Check that the file has permissions to run as an executable. 

First, we will use the get_traj() function to create traj, a dataframe of trajectory points. Here, I have set the output directory to store the generated trajectory files, the met directory to store the meteorological data files, and the executive directory to path to the hyst_std executable file. 
```{r}
# traj <- get_traj(trajDate = '2019-06-01',
#                  OutDir = '/home/donatello/Documents',
#                  MetDir = '/home/donatello/Met',
#                  ExecDir = '/home/donatello/Rprojects/testtraj/')
```

More often, you will generate trajectories for multiple dates at a time. To do so, first create an object, dates, to contain your desired range of dates. Make sure the dates are in character format. 
```{r}
# dates <- seq(as.Date('2019-02-01'),
#              as.Date('2019-02-25'),
#              "days") %>%
#   as.character()
```

Then, using the map_dfr function from the purrr package, run the get_traj() function over each date you listed. Save the output dataframe as an RDS file for future use. 
```{r}
# traj <- map_dfr(dates, ~get_traj(.,
#                                  OutDir = 'home/donatello/Documents',
#                                  MetDir = '/home/donatello/Met',
#                                  ExecDir = '/home/donatello/Rprojects/testtraj'))
# saveRDS(traj, '/home/donatello/Rprojects/traj_example')
```

The dataframe traj now contains the interpolated trajectory points. Using tools from tidyverse, filter the points down to my area of interest. 
```{r}
# read in trajectory data
traj <- readRDS('/home/donatello/Rprojects/traj_example') %>%
  # filter for points within Vallejo boundaries
  filter(lat >= 38.0 &
         lat <= 38.2 &
         lon >= -122.38 &
         lon <= -122.15)
```

Then, using the terra package, create a blank grid to map the trajectories to. 
```{r}
# get extent
trajExt <- cbind(traj$lon, traj$lat) %>%
  ext()
# create blank grid from extent
trajgrid <- rast(trajExt, 
               resolution = 0.005, 
               crs = '+proj=longlat +datum=WGS84')
```

Now, use the rasterize() function from the terra package to create a frequency plot of the trajectory points. See the frequency_plot.png in folder for the frequency plot.  
```{r}
# create coordiates matrix for rasterize function
coords <- cbind(traj$lon, traj$lat)

# generate frequency plot
outputs <- rasterize(coords, trajgrid, fun = 'count')
plot(outputs)
```

